---
title: "R_gam_M1_diff_dfs_and_models_superlet20_prevmodels_logpower_allelbrots_zscored"
format: html
editor: visual
---

# Load data & libs

```{r}
library(mgcv)
library(nlme)
library(abind)
library(gratia)
library(ggeffects)
library(emmeans)
library(tidyverse)
library(itsadug)
```

```{r}
subjects = c('S01', 'S02', 'S03', 'S04', 'S05', 'S06', 'S07', 'S08', 'S09', 'S10', 'S11', 'S12', 'S13', 'S14', 'S15', 'S16')

print('Load data')
df_subj <- data.frame()
for (s in 1:length(subjects)) {
  subj <- subjects[s]
  print(subj)
  df <-read.csv(paste('/run/media/okapi/TOSHIBA EXT/for_r/', subj, '_M1_superlet_20_alpha10_and_12_beta15_and_betavolume_rot_inh_-08to03_allepo_zscored_ind_vert.csv', sep=''))
  df_subj <- rbind(df_subj,df)
}

df <- df_subj
df$Subject <- as.factor(df$Subject)
```

# log-powers, unlog-beta volume + z-scored data + without excluding large elbow rotations

## Visualize data

### Elbow rotation distribution in subjects

```{r}
df %>%
  filter(Rotation < .5) %>%
  distinct(Subject, Trial, Rotation) %>%
  ggplot(aes(x = Rotation, fill = (Rotation<=.05))) +
  facet_wrap(~ Subject, ncol = 4) +
  geom_histogram(breaks = seq(0, .5, .05))
```

### Alpha Power 10 Hz distribution in subjects

```{r}
df %>%
  filter(Rotation < .5) %>%
  ggplot(aes(x = AlphaPower10)) +
  facet_wrap(~ Subject, ncol = 4) +
  geom_histogram(bins=50)
```

### Alpha Power 12 Hz distribution in subjects

```{r}
df %>%
  filter(Rotation < .5) %>%
  ggplot(aes(x = AlphaPower12)) +
  facet_wrap(~ Subject, ncol = 4) +
  geom_histogram(bins=50)
```

### Beta Power 15 Hz distribution in subjects

```{r}
df %>%
  filter(Rotation < .5) %>%
  ggplot(aes(x = BetaPower15)) +
  facet_wrap(~ Subject, ncol = 4) +
  geom_histogram(bins=50)
```

### Beta volume distribution in subjects

```{r}
df %>%
  filter(Rotation < .5) %>%
  ggplot(aes(x = BetaVolume)) +
  facet_wrap(~ Subject, ncol = 4) +
  geom_histogram(bins=50)
```

### Elbow rotation overall, histogram:

```{r}
hist(df$Rotation, breaks = 50)
```

### Alpha Power 10Hz overall, histogram:

```{r}
hist(df$AlphaPower10, breaks = 50)
```

### Alpha Power 12Hz overall, histogram:

```{r}
hist(df$AlphaPower12, breaks = 50)
```

### Beta Power 15Hz overall, histogram:

```{r}
hist(df$BetaPower15, breaks = 50)
```

### Beta Volume overall, histogram:

```{r}
hist(df$BetaVolume, breaks=50)
```

## Run models + visualize predictions

### Alpha 10Hz only + trials: predict Elbow Rotation

```{r}
model_a10 = bam(Rotation ~ s(Trial, Subject, bs='re') +te(AlphaPower10,Time, k=5),
                data=df, method='REML')
```

```{r}
(ggemmeans(model_a10, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower10 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) + geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation') +labs(color = "AlphaPower10")
```

### Alpha 10Hz only + no trials: predict Elbow Rotation

```{r}
model_a10_notri = bam(Rotation ~ te(AlphaPower10,Time,k=5), # + s(Time) +ti(AlphaPower,Time),
                data=df, method='REML')
```

```{r}
(ggemmeans(model_a10_notri, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower10 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation') +labs(color = "AlphaPower10")
```

### Alpha 12Hz only + trials: predict Elbow Rotation

```{r}
model_a12 = bam(Rotation ~ s(Trial, Subject, bs='re') +te(AlphaPower12,Time),
                data=df, method='REML')
```

```{r}
(ggemmeans(model_a12, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower12 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) + geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation') +labs(color = "AlphaPower12")
```

### Alpha 12Hz only + no trials: predict Elbow Rotation

```{r}
model_a12_notri = bam(Rotation ~ te(AlphaPower12,Time), 
                data=df, method='REML')
```

```{r}
(ggemmeans(model_a12_notri, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower12 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation') +labs(color = "AlphaPower12")
```

### Beta 15Hz only + trials: predict Elbow Rotation

```{r}
model_b15 = bam(Rotation ~ s(Trial, Subject, bs='re') +te(BetaPower15,Time,k=5),
                data=df, method='REML')
```

```{r}
(ggemmeans(model_b15, terms = c("Time [-0.8:0.3, by=0.1]", "BetaPower15 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) + geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation') +labs(color = "BetaPower15")
```

### Beta 15Hz only + no trials: predict Elbow Rotation

```{r}
model_b15_notri = bam(Rotation ~ te(BetaPower15,Time), 
                data=df, method='REML')
```

```{r}
(ggemmeans(model_b15_notri, terms = c("Time [-0.8:0.3, by=0.1]", "BetaPower15 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation') +labs(color = "BetaPower15")
```

### Beta volume only + trials: predict Elbow Rotation

```{r}
model_b = bam(Rotation ~ s(Trial, Subject, bs='re') +te(BetaVolume,Time), #s(Time) +ti(BetaVolume,Time),
                data=df, method='REML')
```

```{r}
(ggemmeans(model_b, terms = c("Time [-0.8:0.3, by=0.1]", "BetaVolume [-1.5:3, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "BetaVolume")
```

### Beta volume only + no trials: predict Elbow Rotation

```{r}
model_b_notri = bam(Rotation ~ te(BetaVolume,Time),
                data=df, method='REML')
```

```{r}
(ggemmeans(model_b_notri, terms = c("Time [-0.8:0.3, by=0.1]", "BetaVolume [-1.5:3, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c() + xlab('Time, s') + ylab('Elbow Rotation') +labs(color = "BetaVolume")
```

### Alpha & Beta (all factors: alpha 10Hz, alpha 12Hz, beta 15Hz, beta volume) + trials: predict Elbow Rotation

```{r}
set.seed(1)
model_ab = bam(Rotation ~ s(Trial, Subject, bs='re') +te(AlphaPower10,Time,k=5) +te(AlphaPower12,Time,k=5) + te(BetaVolume,Time,k=5) +te(BetaPower15,Time),
                data=df, method='REML')
```

#### Alpha Power 10Hz prediction + trials

```{r}
(ggemmeans(model_ab, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower10 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "AlphaPower10")
```

#### Alpha Power 12Hz prediction + trials

```{r}
(ggemmeans(model_ab, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower12 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "AlphaPower12")
```

#### Beta Power 15Hz prediction + trials

```{r}
(ggemmeans(model_ab, terms = c("Time [-0.8:0.3, by=0.1]", "BetaPower15 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "BetaPower15")
```

#### Beta Volume prediction + trials

```{r}
(ggemmeans(model_ab, terms = c("Time [-0.8:0.3, by=0.1]", "BetaVolume [-1.5:3, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "BetaVolume")
```

### Alpha & Beta (all factors: alpha 10Hz, alpha 12Hz, beta 15Hz, beta volume) + no trials: predict Elbow Rotation

```{r}
set.seed(1)
model_ab_notri = bam(Rotation ~ te(AlphaPower10,Time) +te(AlphaPower12,Time) + te(BetaVolume,Time,k=5) +te(BetaPower15,Time),
                data=df, method='REML')
```

#### Alpha Power 10Hz prediction + no trials

```{r}
(ggemmeans(model_ab_notri, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower10 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "AlphaPower10")
```

#### Alpha Power 12Hz prediction + no trials

```{r}
(ggemmeans(model_ab_notri, terms = c("Time [-0.8:0.3, by=0.1]", "AlphaPower12 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "AlphaPower12")
```

#### Beta Power 15Hz prediction + no trials

```{r}
(ggemmeans(model_ab_notri, terms = c("Time [-0.8:0.3, by=0.1]", "BetaPower15 [-2:2, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "BetaPower15")
```

#### Beta Volume prediction + no trials

```{r}
(ggemmeans(model_ab_notri, terms = c("Time [-0.8:0.3, by=0.1]", "BetaVolume [-1.5:3, by=0.2]"))) %>% ggplot(aes(x = x, y = predicted, ymin=conf.low, ymax=conf.high, color=as.numeric(as.character(group)), group = group)) +     geom_line() + scale_color_viridis_c()+ xlab('Time, s') + ylab('Elbow Rotation')  +labs(color = "BetaVolume")
```

## Summary of models

```{r}
summary(model_a10)
```

```{r}
summary(model_a12)
```

```{r}
summary(model_a10_notri)
```

```{r}
summary(model_a12_notri)
```

```{r}
summary(model_b15)
```

```{r}
summary(model_b)
```

```{r}
summary(model_b15_notri)
```

```{r}
summary(model_b_notri)
```

```{r}
summary(model_ab)
```

```{r}
summary(model_ab_notri)
```

## AIC model comparison

```{r}
AIC(model_a10) - AIC(model_a10_notri)
```

```{r}
AIC(model_a12) - AIC(model_a12_notri)
```

```{r}
AIC(model_b15) - AIC(model_b15_notri)
```

```{r}
AIC(model_b) - AIC(model_b_notri)
```

```{r}
AIC(model_a12) - AIC(model_a10)
```

```{r}
AIC(model_b15) - AIC(model_a12)
```

```{r}
AIC(model_b15) - AIC(model_b)
```

```{r}
AIC(model_ab) - AIC(model_b15)
```

```{r}
AIC(model_ab) - AIC(model_ab_notri)
```

## Check models

```{r}
gam.check(model_a10)
```

```{r}
gam.check(model_a12)
```

```{r}
gam.check(model_b15)
```

```{r}
gam.check(model_b)
```

```{r}
gam.check(model_a10_notri)
```

```{r}
gam.check(model_b15_notri)
```

```{r}
set.seed(1)
gam.check(model_ab)
```

```{r}
set.seed(1)
gam.check(model_ab_notri)
```
